---
title: "R Notebook"
output: html_notebook
---

## Libraries and import 

```{r}
library(tidyverse)
library(lubridate)
library(httr)
library(WikidataR)
library(svglite)
library(openssl)
```




## Files

Nomenclator - available here:
https://opendata-ajuntament.barcelona.cat/data/es/dataset/nomenclator-bcn
```{r}
openDataNamedStreets <- read_csv("0401230100_BCN_Nomenclator_OpenData.csv")
```

Carrerer - available here:
https://opendata-ajuntament.barcelona.cat/data/es/dataset/carrerer
```{r}
allStreetNames <- read_csv("allStreetNames.csv")
```

Augmented data set - available here:
https://sharingcities.thedata.place/dataset/nou-nomenclator-bcn
```{r}
sharingDataNamedStreets <- read_csv("streetNamesWithWikipediaInfo.csv")
```

Neighborhood data - available here:
https://opendata-ajuntament.barcelona.cat/data/en/dataset/20170706-districtes-barris/
```{r}
barrios <-  read_csv("districtes_i_barris_170705.csv")
```


Census data - available here:
https://opendata-ajuntament.barcelona.cat/data/es/dataset/taula-direle-sec-cens
```{r}
censusAddreses <- read_csv("censusData.csv")
```


## Data tidying

### 1 allStreetNames

* Removed empty columns
* Converted columns to correct type
```{r}
allStreetNamesTidy <- allStreetNames %>%
  transmute(streetCode = as.numeric(CODI_VIA),
         CODI_CARRER_INE = as.numeric(CODI_CARRER_INE),
         streetTypeAb = as.factor(TIPUS_VIA),
         officialStreetName = NOM_OFICIAL,
         shortStreetName = NOM_CURT) 
```



### openDataNamedStreets

* Removed empty columns
* Converted columns to correct type
```{r}
openDataNamedStreetsTidy <- openDataNamedStreets %>%
  transmute(streetCode = as.numeric(CODI),
         streetType = as.factor(TIPUS_VIA),
         PARTICULES = PARTICULES,
         streetNameNomlcator = NOM,
         dateApproved = dmy(DATA_APROV),
         notesOnDate = NOTES_DATA,
         streetNameInfo = DESCRIPCIO1,
         referenceDocument = FONTS_DOC,
         otherStreetNameNomclator = ALTRES_NOMS) 
```




### sharingDataNamedStreets

* Removed empty columns
* Converted columns to correct type
* Not planning to use, but retaining for reference

```{r}
sharingDataNamedStreetsTidy <- sharingDataNamedStreets %>%
  transmute(OBJECTID = OBJECTID,
            CODI_AJB = CODI_AJB,
            CODI_INE = CODI_INE,
            GRUP_TVIA = GRUP_TVIA,
            streetTypeCodeAb = as.factor(T_Via),
            streetType = as.factor(Desc_Via),
            streetNameNomclatorShort = N_Carrer,
            streetNameShort = N_CarrerC,
            streetName = N_CarrerC2,
            StreetNameNomclator = NOM_indexOnomastic,
            dateApproved = dmy(DATA_APROV),
            notesOnDate = NOTES_DATA,
            gender = as.factor(ParaulaClau_2),
            profession = as.factor(ParaulaClau_3),
            wikipediaPage = PAGINA_WIKIPEDIA,
            biography = BIOGRAFIA_CAT,
            wikipediaPage2 = PAGINA_WIKIPEDIA2)
```



### barrio

* only tidied up column names
```{r}
barriosTidy <- barrios %>%
  select(districtCode = CODI_DISTRICTE,
         districtName = NOM_DISTRICTE,
         neighborhoodCode = CODI_BARRI,
         neighborhoodName = NOM_BARRI)
```


### censusAddreses

* Converted columns to correct type
```{r}
censusAddresesTidy <- censusAddreses %>%
  transmute(streetCode = as.numeric(CARRER),
            streetName = NOM_CARRER,
            doorNumberFirst = as.numeric(PRIMER_NUM),
            doorNumberLast = as.numeric(ULTIM_NUM),
            TIPUSNUM = TIPUSNUM,
            districtCode = as.numeric(DISTRICTE),
            SECC_CENS = SECC_CENS,
            AEB = as.numeric(AEB),
            neighborhoodCode = as.numeric(BARRI),
            SECC_EST = SECC_EST,
            postcode = as.factor(DPOSTAL))
            
```

* Making a subsetted dataset with only the useful lookup columns
```{r}
censusNeighborhoodLookup <- censusAddresesTidy %>%
  select(streetCode, districtCode, neighborhoodCode, postcode) %>%
  distinct()
```



## Joining tidydata sets


* Some duplicates exist for streets that passthrough multiple neighborhoods
* Converted neighborhoods and districts to list
```{r}
streetNamesWithNeighborhoods <- allStreetNamesTidy %>%
  left_join(censusNeighborhoodLookup, by = "streetCode") %>%
  select(streetCode, officialStreetName, districtCode, neighborhoodCode) %>%
  distinct() %>%
  left_join(barriosTidy, by = "neighborhoodCode") %>%
  select(streetCode, officialStreetName, districtName, neighborhoodName) %>%
  nest(-streetCode, -officialStreetName)
  
```


* TO-DO i think this is missing because there was no census data
```{r}
(missingSteetData <- streetNamesWithNeighborhoods %>%
  mutate(joinTest = "streetnames") %>%
  full_join(openDataNamedStreetsTidy, by = "streetCode") %>%
  filter(is.na(joinTest)))
```


```{r}
streetsNeighborhoodNomclator <- streetNamesWithNeighborhoods %>%
  full_join(openDataNamedStreetsTidy, by = "streetCode")
```


## Retrieving raw itemCodes on wikidata, genderCodes and year of birth

### Getting gender items from street name Nomclator

```{r}
nomclatorForItemLookup <- streetsNeighborhoodNomclator %>%
  select(streetCode, streetNameNomlcator)
```

Total number of items to look up in wikidata

```{r}
nrow(nomclatorForItemLookup)
```

Retrieving wikidata item codes

```{r}
getItemCodesNomclator <- nomclatorForItemLookup %>%
  mutate(itemToFind = streetNameNomlcator) %>%
  nest(itemToFind) %>%
  mutate(itemSearch = map(data, ~find_item(.)),
         itemCodeToSearch = map(itemSearch, ~pluck(., 1)),
         itemCode = map(itemCodeToSearch, ~pluck(., "id"))) %>%
  unnest(itemCode) 
```

Creating a simplified table of item codes to look up

```{r}
itemCodesToGetPropertiesNomclator <- getItemCodesNomclator %>%
  select(streetCode, itemCode)
```


Retrieving properties of assosiated items

```{r}
pluckedGenderIdByItemCodeNomclator <- itemCodesToGetPropertiesNomclator %>%
  mutate(wikiDataObject = itemCode,
         wikiDataObject = map(wikiDataObject, ~get_item(.)), 
         masterObject = map(wikiDataObject, ~pluck(., 1)),
         claimsObject = map(masterObject, ~pluck(.,"claims")),
         genderObject = map(claimsObject, ~pluck(.,"P21")),
         genderMainsnakObject = map(genderObject, ~pluck(.,"mainsnak")),
         genderMainsnakDvObject = map(genderMainsnakObject, ~pluck(.,"datavalue")),
         genderMainsnakDvValObject = map(genderMainsnakDvObject, ~pluck(.,"value")),
         genderId = map(genderMainsnakDvValObject, ~pluck(., "id"))) %>%
  unnest(genderId)%>%
  select(streetCode, itemCode, genderId)
```


```{r}
pluckedGenderIdByItemCodeDistinctNomclator <- pluckedGenderIdByItemCodeNomclator %>%
  distinct() 
```


Number of matches Nomlator
```{r}
nrow(pluckedGenderIdByItemCodeDistinctNomclator)
```





### Getting item codes and genders from description via regex

```{r}
experiment <- streetsNeighborhoodNomclator %>%
  select(streetCode, streetNameInfo, streetNameNomlcator) %>%
  mutate(verificationNameFromInfoBeforeBracket = str_extract(streetNameInfo, "[a-zA-ZÀÁÇÉÈÍÏÓÒÚÜ·àáçéèíïóòñúü'\\-\\s]+(?=\\()"),
         verificationNameFromInfoBeforeFirstDot = str_extract(streetNameInfo, "[a-zA-ZÀÁÇÉÈÍÏÓÒÚÜ·àáçéèíïóòñúü'\\-\\s]+(?=\\.|\\,)"),
         streetNameShort = str_trunc(streetNameInfo, width = 80)) %>%
  select(streetCode, verificationNameFromInfoBeforeBracket, verificationNameFromInfoBeforeFirstDot) %>%
  pivot_longer(cols = c(verificationNameFromInfoBeforeBracket,verificationNameFromInfoBeforeFirstDot)) %>%
  filter(!is.na(value))
```

```{r}
experimentGetItemCodes <- experiment %>%
  mutate(itemToFind = value) %>%
  nest(itemToFind) %>%
  mutate(itemSearch = map(data, ~find_item(.)),
         itemCodeToSearch = map(itemSearch, ~pluck(., 1)),
         itemCode = map(itemCodeToSearch, ~pluck(., "id"))) %>%
  unnest(itemCode) 
```


```{r}
experimentGetItemCodesToGetProperties <- experimentGetItemCodes %>%
  select(streetCode, itemCode)
```

```{r}
experimentPluckedGenderIdByItemCodeRegex <- experimentGetItemCodesToGetProperties %>%
  mutate(wikiDataObject = itemCode,
         wikiDataObject = map(wikiDataObject, ~get_item(.)), 
         masterObject = map(wikiDataObject, ~pluck(., 1)),
         claimsObject = map(masterObject, ~pluck(.,"claims")),
         genderObject = map(claimsObject, ~pluck(.,"P21")),
         genderMainsnakObject = map(genderObject, ~pluck(.,"mainsnak")),
         genderMainsnakDvObject = map(genderMainsnakObject, ~pluck(.,"datavalue")),
         genderMainsnakDvValObject = map(genderMainsnakDvObject, ~pluck(.,"value")),
         genderId = map(genderMainsnakDvValObject, ~pluck(., "id"))) %>%
  unnest(genderId)%>%
  select(streetCode, itemCode, genderId)
```


```{r}
experimentPluckedGenderIdByItemCodeRegexExperiment <- experimentPluckedGenderIdByItemCodeRegex %>%
  distinct() 
```


Number of matches Nomlator
```{r}
nrow(experimentPluckedGenderIdByItemCodeRegexExperiment)
```




### Merging itemCode sets


```{r}
nomclatorAndRegexMerged <- bind_rows(experimentPluckedGenderIdByItemCodeRegexExperiment, pluckedGenderIdByItemCodeDistinctNomclator)
```
```{r}
nomclatorAndRegexMerged %>%
  distinct() 
```



### Discarding manually incorrect items where a duplicate was found (~60 streets)

```{r}
itemsToDiscard <- c("Q261691", "Q61630564", "Q48900", "Q319439","Q19998062", "Q50464207", "Q8962231", "Q759804", "Q21625441", "Q22103226","Q83321", "Q11915106", "Q5287319", "Q5833377", "Q11921849","Q60826742","Q21512998","Q5864726","Q5873187","Q5931452", "Q29253032", "Q21832771", "Q55897295","Q18719870","Q18379905", "Q1345186", "Q1706315", "Q705946", "Q6751540", "Q11934916", "Q31966", "Q16161344", "Q19458190", "Q267858", "Q1771132", "Q9033484", "Q231156", "Q55068526", "Q2406301", "Q4339403", "Q29253032","Q25400539", "Q356258", "Q16179820", "Q150575", "Q11932279", "Q11704317","Q21289900", "Q38589936", "Q20658483", "Q163543", "Q60827925", "Q64026936", "Q40365895", "Q66774041", "Q164294", "Q10307409", "Q5931538", "Q649731", "Q19291019", "Q17301468", "Q96359163", "Q45426155", "Q34189", "Q11932279", "Q2290455", "Q234253", "Q20101662", "Q5350072")
```

```{r}
`%notin%` <- Negate(`%in%`)
```


```{r}
nomclatorAndRegexMergedDeduplicated <- nomclatorAndRegexMerged %>%
  filter(itemCode %notin% itemsToDiscard)
```


### Recoding wikidata gender codes

```{r}
nomclatorAndRegexTidy <- nomclatorAndRegexMergedDeduplicated %>%
  distinct() %>%
  mutate(genderLabel = as.factor(genderId),
         genderLabel = fct_recode(genderLabel, Female = "Q6581072", Male = "Q6581097")) %>%
  select(streetCode, itemCode, genderLabel)
```



### Retrieving birth years


```{r}
birthYearByItemCode <- nomclatorAndRegexTidy %>%
  mutate(wikiDataObject = itemCode,
         wikiDataObject = map(wikiDataObject, ~get_item(.)),
         masterObject = map(wikiDataObject, ~pluck(., 1)),
         claimsObject = map(masterObject, ~pluck(.,"claims")),
         yearOfBirthObject = map(claimsObject, ~pluck(.,"P569")), 
         yearOfBirthMainSnakObject = map(yearOfBirthObject, ~pluck(.,"mainsnak")),
         yearOfBirthMainsnakDvObject = map(yearOfBirthMainSnakObject, ~pluck(.,"datavalue")),
         yearOfBirthMainsnakDvValObject = map(yearOfBirthMainsnakDvObject, ~pluck(.,"value")),
         yearOfBirthTimestamp = map(yearOfBirthMainsnakDvValObject, ~pluck(.,"time"))) %>%
  unnest(yearOfBirthTimestamp) %>%
  mutate(yearOfBirthYearDirty = str_remove(yearOfBirthTimestamp, "(-.*)"),
         yearOfBirthYearDirty = str_remove(yearOfBirthYearDirty, "\\+"),
         yearOfBirth = as.numeric(yearOfBirthYearDirty)) %>%
  group_by(itemCode) %>%
  summarise(min(yearOfBirth))

```


```{r}
itemCodesWithGenderAndBirthYear <- nomclatorAndRegexTidy %>%
  left_join(birthYearByItemCode, by = "itemCode") %>%
  mutate(yearOfBirth = `min(yearOfBirth)`) %>%
  select(-`min(yearOfBirth)`)
```


## Joining datasets together

```{r}
nomclatorWithGenderAndBirthYear <- streetsNeighborhoodNomclator %>%
  left_join(itemCodesWithGenderAndBirthYear, by = "streetCode")
```



## visualisations - input to infographic

```{r}
nomclatorWithGenderAndBirthYear %>%
  summarise(n_distinct(streetCode))
```




### Overall current proportions where a human with a gender was matched
```{r}
nomclatorWithGenderAndBirthYear %>%
  count(genderLabel) %>%
  filter(!is.na(genderLabel)) %>%
  mutate(proportion = round((n / sum(n)*100), digits = 2))
```


### Histogram of streets by named year with output to svg

```{r}
fiveYearHistogram
```


```{r}
ggsave(file="fiveYearHistogram.svg", plot=fiveYearHistogram, width=10, height=8)
```


### Predictions to 2030
```{r}
newStreetsBy2030Estimate <- nomclatorWithGenderAndBirthYear %>%
  mutate(yearApproved = year(dateApproved)) %>%
  count(yearApproved) %>%
  filter(!is.na(yearApproved), yearApproved >= 2010) %>%
  summarize(mean = mean(n)) %>%
  mutate(prediction = mean * 10) %>%
  select(prediction) %>%
  pull()
```

```{r}
totalIdentifiedFemaleStreets <- nomclatorWithGenderAndBirthYear %>%
  mutate(yearApproved = year(dateApproved)) %>%
  mutate(genderLabelChart = if_else(is.na(genderLabel), "No data or not a person", as.character(genderLabel))) %>%
  filter(genderLabelChart == "Female") %>%
  count() %>%
  pull()
```


```{r}
totalIdentifiedNamedStreets <- nomclatorWithGenderAndBirthYear %>%
  mutate(yearApproved = year(dateApproved)) %>%
  mutate(genderLabelChart = if_else(is.na(genderLabel), "No data or not a person", as.character(genderLabel))) %>%
  filter(genderLabelChart != "No data or not a person") %>%
  count() %>%
  pull()
```


```{r}
(newStreetsBy2030Estimate + totalIdentifiedFemaleStreets) / (totalIdentifiedNamedStreets + newStreetsBy2030Estimate)
```



### Neighborhoods by male/female/no match split

```{r}
biggestNeighborhoods <- nomclatorWithGenderAndBirthYear %>%
  unnest(data) %>%
  count(neighborhoodName) %>%
  filter(n >= 50)



nomclatorWithGenderAndBirthYear %>%
  unnest(data) %>%
  count(neighborhoodName, genderLabel) %>%
  semi_join(biggestNeighborhoods, by = "neighborhoodName") %>%
  ggplot(aes(neighborhoodName, n, fill = genderLabel)) +
  geom_col(position = "fill") +
  coord_flip()
```


## Downloading images

```{r}
retrievingImageNames <- nomclatorWithGenderAndBirthYear %>%
  filter(genderLabel == "Female") %>%
  mutate(wikiDataObject = itemCode,
         wikiDataObject = map(wikiDataObject, ~get_item(.)),
         masterObject = map(wikiDataObject, ~pluck(., 1)),
         claimsObject = map(masterObject, ~pluck(.,"claims")),
         imageObject = map(claimsObject, ~pluck(.,"P18")), 
         imageObjectMainSnakObject = map(imageObject, ~pluck(.,"mainsnak")),
         imageObjectMainsnakDvObject = map(imageObjectMainSnakObject, ~pluck(.,"datavalue")),
         imageObjectMainsnakDvValObject = map(imageObjectMainsnakDvObject, ~pluck(.,"value"))) %>%
  unnest(imageObjectMainsnakDvValObject) %>%
  select(itemCode, streetNameNomlcator, imageObjectMainsnakDvValObject)
```



```{r}
imagesToDownload <- retrievingImageNames %>%
  mutate(imageObjectMainsnakDvValObjectUnderscore = str_replace_all(imageObjectMainsnakDvValObject, " ", "_"),
         imageNameMd5 = md5(imageObjectMainsnakDvValObjectUnderscore),
         a = substr(imageNameMd5, 0,1),
         ab = substr(imageNameMd5, 1,2),
         imageDownloadUrl = str_c("https://upload.wikimedia.org/wikipedia/commons/", a, "/", ab, "/", imageObjectMainsnakDvValObjectUnderscore),
         imageName = str_c("images/", streetNameNomlcator, str_extract(imageObjectMainsnakDvValObject, "(\\.)(.*)")))
```

```{r}
imageUrl <- imagesToDownload %>%
  select(imageDownloadUrl) %>%
  pull()
```

```{r}
imageName <- imagesToDownload %>%
  select(imageName) %>%
  pull()
```

```{r}
download.file(url=imageUrl, destfile = imageName, mode="wb")
```


## Exporting csv for submission

```{r}
outputCsv <- nomclatorWithGenderAndBirthYear %>%
  unnest(data) %>%
  select(-PARTICULES)
```

```{r}
write_csv(outputCsv, "outputCsv.csv")
```




